---
title: "climplot: Streamline Plotting the Walter & Lieth Climatic Diagram"
pagetitle: "Introduction to climplot"
author: 
- <a href="https://gitee.com/WYY_Space"> Auman Chan, PhD Candidate</a>
date:  "`r Sys.Date()`"
output: html_document
    
vignette: >
  %\VignetteIndexEntry{1. Introduction and Intro of climplot}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
  editor_options: 
  chunk_output_type: console
params:
  EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
---

```{css,echo=FALSE}

body {
  margin-top: 16px;
  margin-bottom: 16px;
  font-size: 19px;
  font-weight: 200;
}

#sidebar {
font-size: 16px
}

#sidebar p {
font-size: 16px
}

.contents h1, .contents h2, .contents h3, .contents h4 {
  padding-top: 60px;
  margin-top: -50px;
}

h4 {
  font-size: 19px;
}

h3 {
  font-size: 22px;
}

h2 {
  font-size: 26px;
}

h1 {
  font-size: 30px;
}

pre {
  border: 0;
  overflow-x: auto;
  white-space: pre-wrap;
  white-space: -moz-pre-wrap;
  white-space: -pre-wrap;
  white-space: -o-pre-wrap;
  word-wrap: break-word;
  background: #e9f7ed;
    color: #ffffff;
    padding-top: 0.5em;
  padding-bottom: 0.5em;
  padding-left: 20px;
  padding-right: 20px;
  line-height: 21px;
  font-size: 15px;
}

pre code {
  color: #182125;
}

code {
  word-wrap: break-word;
  font-size: 85%;
  font-family: Consolas,Monaco,"Courier New",monospace;
  border-radius: 0;
}

.nav-pills>li>a {
  border-radius: 0;
}

.navbar-default .navbar-nav>.active>a, .navbar-default .navbar-nav>.active>a:focus, .navbar-default .navbar-nav>.active>a:hover {
  color: #408AB4;
}
.navbar-default .navbar-nav>li>a {
  color: #408AB4;
}

table p {
  font-size: 85%;
}

.ref-index h2 {
  font-size: 24px;
}

h1.title {
  font-weight: bold;
  font-size: 180%;
  text-align: center
}

h2.subtitle {
  font-size: 120%;
  text-align: center
}

h3.subtitle {
  font-size: 120%;
  text-align: center
}

h4.author {
  font-size: 120%;
  text-align: center
}

h4.date {
  font-size: 120%;
  text-align: center
}

p {
  margin-bottom: 30px;
}
```

[climplot](https://gitee.com/WYY_Space/climplot) is intended to to acquire essential climate data of Walter & Lieth climatic diagram of locations worldwide, and plot the climatic diagram in a more convenient and customized way.

<div class="alert alert-info">

The main features of the package are: 

- The use of worldwide climate data for acquiring standard and reliable data for
Walter & Lieth climatic diagram plotting
- The provision for more parameter to customize the plots and display more information



The package offers functions for:

- Acquire climate data to plot Walter & Lieth Climatic Diagram for locations worldwide
- Plot the Walter & Lieth climatic diagram 
- Modify the color and information display on the diagram

</div>

# Installing and loading

To install the latest developmental version from [gitee](https://gitee.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes)and [git2r](https://cran.r-project.org/package=git2r):

```{r intasll, eval=FALSE}
install.packages('remotes')
install.packages('git2r')
remotes::install_git("https://gitee.com/WYY_Space/climplot.git")
```

```{r include=FALSE}
#load package
library(climplot)
```

# Climate data extraction

Finding and clearing up climate data of numerous location is a tough work. [climplot](https://gitee.com/WYY_Space/climplot) can acquire data from 
Worldclim Historical monthly weather data in 2010-2019(Version of 2.5 minutes, https://worldclim.org/data/monthlywth.html), and arrange them to be ready for diagram plotting.

## Data preparation



```{r global options, echo=FALSE, message=FALSE, warning=FALSE}

data("locdata")
locdata
```

